---
title: "Augustine-PS2.qmd."
author: "Christopher Augustine"
format: html
editor: visual
---

setting things up 
```{r}
rm(list=ls())
library(tidyverse)
library(here)
```
reading in the data
```{r}
halloween<-read.csv(here("data", "Halloween_data_PS2.csv"))
```

using 2 function to examine that data
```{r}
glimpse(halloween)
summary(halloween)
```
fixing the data 
```{r}
halloween<- halloween %>%
  mutate(
  costume=as.factor(costume), 
  Costume_type=as.factor(Costume_type),
  sugar_high=as.factor(sugar_high))
```

```{r}
levels(halloween$costume)
levels(halloween$sugar_high)
levels(halloween$Costume_type)
```
fixing the differences in names in data
```{r}
fix_witch <- which(halloween$costume %in% c("witch", "Witch", "witch "))
halloween$costume[fix_witch] <- "witch"

fix_zombie <- which(halloween$costume %in% c("zombie", "Zombie"))
halloween$costume[fix_zombie] <- "zombie"

fix_unicorn <- which(halloween$costume %in% c("unicorn", "Unicorn"))
halloween$costume[fix_unicorn] <- "unicorn"

fix_not_scary <- which(halloween$Costume_type %in% c("not scary", "Not scary", "notscary"))
halloween$Costume_type[fix_not_scary] <- "not scary"

fix_scary <- which(halloween$Costume_type %in% c("scaly","scary","Scary","scary "))
halloween$Costume_type[fix_scary] <- "scary"

fix_no <- which(halloween$sugar_high %in% c("no", "No", "NO"))
halloween$sugar_high[fix_no] <- "no"

fix_yes <- which(halloween$sugar_high %in% c("yes", "Yes"))
halloween$sugar_high[fix_yes] <- "yes"

#to update levels
halloween$costume<- factor(halloween$costume)
halloween$Costume_type<- factor(halloween$Costume_type)
halloween$sugar_high<- factor(halloween$sugar_high)
```

task #7 making a scatterplot to look like the one provided, using min_temp_trick_treat_C to show temperature and viewing that across age 
```{r}
ggplot(halloween, aes(x = age, y = min_temp_trick_treat_C)) +
  geom_point(aes(color = Costume_type, shape = Costume_type), size = 3, alpha = 0.7)  +  
  labs(
    x = "Age of trick-or-treater (years)",
    y = "Minimum temperature while out (C)",
  ) +
  theme_bw() 
```
task #8 filtering out 10 year old and older who ate 100 g or more of candy and making a histogram to match the one provided 

```{r}
candygoblins <- halloween %>%
  filter(age >= 10, g_candy_consumed >= 100)
```


```{r}
ggplot(candygoblins, aes(x = candy_pieces)) +
  geom_histogram(binwidth = .5, alpha = 1.0) +
  labs(
    x = "Number of candy pieces eaten",
    y = "Count"
  ) +
  theme_minimal() 
```
task 9:
summary statics of age and grams of candy consumed showing standard deviation and means then viewing in another tab 
```{r}
halloween %>%
  group_by(sugar_high) %>%
  summarize(
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE),
    mean_g_candy = mean(g_candy_consumed, na.rm = TRUE),
    sd_grams_candy = sd(g_candy_consumed, na.rm = TRUE)
  ) %>%
  View()
```

task 10 
A box and whisker plot showing amount of candy consumed by children with scary, not scary, and uncertain costume types and wether that child got a sugar high from candy eaten
```{r}
ggplot(halloween, aes(x = Costume_type, y = g_candy_consumed)) +
  geom_boxplot(alpha = 0.7, outlier.color = "orange")+
    geom_jitter(color = "orange", width = 0.0, size = 2.5, alpha = 0.8)+
  labs(
    x = "Costume Type",
    y = "Amount of candy consumed (g)"
  ) +
  facet_wrap(~ sugar_high)
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),  
    legend.position = "none"  
    #i added this because for some reason it was adding a legend that was messing other things up, guessing its a settings issue but it made the graph look like the one in the document
  )
```

task 13
I will be the yellow Teletubbie with 3 of my friends so not scary.    
