---
title: "Gardner PS2"
format: html
editor: visual
---

## Problem Set 2 

Normal Set up

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

Read in the data

```{r}
halloween <- read.csv(here("Data", "Halloween_data_PS2.csv"))
```

Examine the data

using `str()`

```{r}
str(halloween)
```

using `glimpse()`

```{r}
glimpse(halloween)
```

Turn `costume_type` , `costume` and `sugar_high`, into factor data

```{r}
halloween$Costume_type <- as.factor(halloween$Costume_type)
levels(halloween$Costume_type) 

halloween$sugar_high <- as.factor(halloween$sugar_high)
levels(halloween$sugar_high)

```

Fix errors in data

fix `Costume_type`

```{r}
fix <- which(halloween$Costume_type %in% c("not scary","Not scary", "notscary"))

halloween$Costume_type[fix] <- "not scary"

halloween$Costume_type <- factor(halloween$Costume_type)
```

```{r}
fix2 <- which(halloween$Costume_type %in% c("scaly", "scary", "Scary", "scary "))

halloween$Costume_type[fix2] <- "scary"

halloween$Costume_type <- factor(halloween$Costume_type)
levels(halloween$Costume_type)
```

Fix `sugar_high`

```{r}
fix3 <- which(halloween$sugar_high %in% c("yes", "Yes"))

halloween$sugar_high[fix3] <- "yes"

halloween$sugar_high <- factor(halloween$sugar_high)
levels(halloween$sugar_high)
```

```{r}
fix <- which(halloween$sugar_high %in% c("NO", "no", "No"))

halloween$sugar_high[fix] <- "no"

halloween$sugar_high <- factor(halloween$sugar_high)
levels(halloween$sugar_high)
```

Make a scatter plot of `min_temp_trick_treat_C` vs `age`

```{r}
ggplot(halloween, aes(x = age, y = min_temp_trick_treat_C)) +
  geom_point(aes(color = Costume_type, shape = Costume_type), size = 5, show.legend = T) + 
  xlab("Age of trick-or-treaters (years)") + 
  ylab("Minimum temperature while out (C)") + 
  theme_bw()
```

Make a new data set with children under the age of 10 and consumed 100 or more grams of candy

```{r}
young_candy <- halloween |> 
  filter(age >= 10, g_candy_consumed >= 100)
```

Make a histogram of how much candy the kids ate

```{r}
ggplot(young_candy, aes(x = candy_pieces)) + 
  geom_histogram( bins = 30) + 
  xlab("Number of candy pieces eaten") + 
  ylab("Count")
```

Summary statistics for the grams of candy eaten and age of children that experienced a sugar high

```{r}
halloween |>
  group_by(sugar_high) |>
  summarize(mean_age = mean(age), sd_age = sd(age),
            mean_grams = mean(g_candy_consumed), sd_grams = sd(g_candy_consumed)) |>
  View()
```

Make a box plot for costume type vs grams of candy consumed separated by whether or not there was a sugar high

```{r}
ggplot(halloween, aes(x = Costume_type, y = g_candy_consumed)) + 
  geom_boxplot() + 
  geom_point(color = "orange") + 
  facet_wrap(~sugar_high) +
  xlab("Costume Type") + 
  ylab("Amount of candy consumed (g)") +
  theme_bw()
```
