---
title: "Kotula-PS2"
format: html
---

### Set Up

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

read in the data and examine using 2 functions

```{r}
Halloween <- read.csv(here("Data", "Halloween_data_PS2.csv"))

str(Halloween)
glimpse(Halloween)
```

### Assign appropriate data types for the data

Costume_type should be a factor with 3 levels: Not scary, Uncertain, and Scary

sugar_high should be a factor with 2 levels: Yes and No

```{r}
Halloween$Costume_type <- as.factor(Halloween$Costume_type)
levels(Halloween$Costume_type) 

Halloween$sugar_high <- as.factor(Halloween$sugar_high)
levels(Halloween$sugar_high)
```

### Fix the factor levels and data entry inconsistencies

Now both variables are factors, but there are too many levels so they must be fixed.

```{r}
fix <- which(Halloween$Costume_type %in% c("not scary",
                                           "Not scary",
                                           "notscary"))

Halloween$Costume_type[fix] <- "Not scary"
Halloween$Costume_type <- factor(Halloween$Costume_type)

fix2 <- which(Halloween$Costume_type %in% c("scaly",
                                           "scary",
                                           "Scary",
                                           "scary "))

Halloween$Costume_type[fix2] <- "Scary"
Halloween$Costume_type <- factor(Halloween$Costume_type)

levels(Halloween$Costume_type) #now Costume_type has 3 levels

fix3 <- which(Halloween$sugar_high %in% c("no",
                                          "No",
                                          "NO"))

Halloween$sugar_high[fix3] <- "No"
Halloween$sugar_high <- factor(Halloween$sugar_high)

fix4 <- which(Halloween$sugar_high %in% c("yes", "Yes"))
Halloween$sugar_high[fix4] <- "Yes"
Halloween$sugar_high <- factor(Halloween$sugar_high)

levels(Halloween$sugar_high) #Now sugar_high has 2 levels
```

The costume variable also has some naming inconsistencies so that needs fixed too.

```{r}
fix5 <- which(Halloween$costume == "Princess")
Halloween$costume[fix5] <- "princess"

fix6 <- which(Halloween$costume == "Unicorn")
Halloween$costume[fix6] <- "unicorn"

fix7 <- which(Halloween$costume == "Witch")
Halloween$costume[fix7] <- "witch"

fix8 <- which(Halloween$costume == "Zombie")
Halloween$costume[fix8] <- "zombie"
```

now all repeat costumes have consistently capitalized names

### Make a scatterplot of min_temp_trick_treat_C vs. age

```{r}
ggplot(Halloween, aes(x = age,
                      y = min_temp_trick_treat_C,
                      color = Costume_type,
                      shape = Costume_type))+
  geom_point(size = 4)+
  xlab("Age of trick-or-treater (years)")+
  ylab("Minimum temperature while out (C)")+
  theme_bw()
```

### Recreate the second plot

Isolate data for children who are 10+ and consumed 100g+ of Halloween candy

```{r}
#isolate the data
big_eaters <- Halloween |>
  filter(age >= 10, g_candy_consumed >= 100)
```

Plot a histogram showing the a count of the number of kids who each ate certain numbers of pieces of candy.

```{r}
ggplot(data = big_eaters, aes(x = candy_pieces))+
  geom_histogram()+
  xlab("Number of candy pieces eaten")+
  ylab("Count")
```

### Summary Statistics for age and g_candy_consumed

```{r}
Halloween |>
  group_by(sugar_high) |>
  summarise(mean_age = mean(age),
            sd_age = sd(age),
            mean_g_candy_consumed = mean(g_candy_consumed),
            sd_g_candy_consumed = sd(g_candy_consumed)) |>
  View()
```

### Recreate the boxplot 

```{r}
ggplot(data = Halloween, aes(x = Costume_type,
                             y = g_candy_consumed))+
  geom_boxplot()+
  geom_point(color = "orange", alpha = 0.7)+
  facet_wrap(~sugar_high)+
  xlab("Costume type")+
  ylab("Amount of candy consumed (g)")+
  theme_bw()
  
```
